---
title: "Functionalizing NEON rasters"
author: "Robert Paul"
date: "June 21, 2016"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)
1. For the CHM, set values == 0 to NA (not trees)
1. Classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that <2m is generally grasses / understory, <6m small trees,and the rest are tall trees. A function could import the desired thresholds. Visualize histogram/density and plot vertical cutoff lines.
1. PLOT - layer the classified raster on top of the hillshade, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)
1. Export the plot figure to a pdf – publishable
1. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.

## Objective 1: Import a Raster

### Load packages
```{r load-packages}
# Load up the packages we need to work with raster data
library(raster)
library(rgdal)
library(rhdf5)
```

### Grab raster
```{r grab-raster}
# Get canopy height model for Teakettle site
chm <- raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")
```

## Objective 2: Visualize distribution of the input data
```{r original-raster-hist}

```

## Objective 3: Set NA values
```{r insert-nas}
# 0 values in the CHM are NA
chm[chm == 0] <- NA
```

## Objective 4: Classify short trees (<6m and >2m)
```{r reclass-trees}

```

## Objective 5: Plot
```{r plot-reclass}

```


## Objective 6: Output to PDF
```{r pdf-output}

```


## Objective 7: Output to GeoTIFF
```{r geotiff-output}

```


